---
title: "ICE 4"
date: "September 9, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  freeze: true
---
#SpatioTemporalPointPatternAnalysis

A spatio-temporal point process is a random collection of points, where each point represents the time and location of an event.



## Install/Launch R

```{R}

```
forest fire or something

```{r}
pacman::p_load(sf, raster, spatstat, sparr, tmap, tidyverse, gifski, animation)
```

## Import the data
```{r}
kbb <- st_read(dsn = "data/rawdata", layer = "Kepulauan_Bangka_Belitung")
```


```{r}
kbb_sf <- st_read(dsn = "data/rawdata", layer = "Kepulauan_Bangka_Belitung") %>%
  st_union() %>%
  st_zm(drop = TRUE, what = "ZM") %>%
  st_transform(crs = 32748)
```


convert kbb to owin object
```{r}
kbb_owin <- as.owin(kbb_sf)
kbb_owin
```
check owin
```{r}
class(kbb_owin)
```

Forest Fire data
```{r}
fire_sf <- read.csv("data/rawdata/forestfires.csv") %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform(crs = 32748)
```

```{r}
fire_sf <- fire_sf %>%
  mutate(DayOfYear = yday(acq_date)) %>%
  mutate(Month_num = month(acq_date)) %>%
  mutate(Month_fac = month(acq_date, label = TRUE, abbr = FALSE))
```

Plotting map showing the distribution of fire points
```{r}
tm_shape(kbb_sf) + tm_polygons() + tm_shape(fire_sf) + tm_dots()
```

Plotting map showing the distribution of fire points by month
```{r}
tm_shape(kbb_sf) + tm_polygons() + tm_shape(fire_sf) + tm_dots(size = 0.1) + tm_facets(by = "Month_fac", free.coords = FALSE, drop.units = TRUE) 
```

##Compute STKDE
###By Month
```{r}
fire_month <- fire_sf %>%
  select(Month_num)
```


make ppp object
```{r}
fire_month_ppp <- as.ppp(fire_month)
fire_month_ppp
```


check ppp object
```{r}
summary(fire_month_ppp)
```

check for duplicated points
```{r}
any(duplicated(fire_month_ppp))
```

combine ppp w owin
```{r}
fire_month_owin <- fire_month_ppp[kbb_owin]
summary(fire_month_owin)
```

```{r}
st_kde <- spattemp.density(fire_month_owin)
summary(st_kde)
```

plotting STKDE by last 6 months 
```{r}
#| fig-width: 12
#| fig-height: 10

tims <- c(7,8,9,10,11,12)
par(mfcol = c(2,3))
for(i in tims){
  plot(st_kde, i,
       override.par = FALSE,
       fix.range = TRUE,
       main = paste("KDE at month", i))
}
```
###By Day
create ppp object
```{r}
fire_yday_ppp <- fire_sf %>%
  select(DayOfYear) %>%
  as.ppp()
```
combine ppp and owin
```{r}
fire_yday_owin <- fire_yday_ppp[kbb_owin]
summary(fire_yday_owin)
```

```{r}
kde_yday <- spattemp.density(fire_yday_owin)
summary(kde_yday)
```

```{r}
#plot(kde_yday)
```

```{r}
library(animation)
library(spatstat)

data(fmd)
fmd_case <- fmd$cases
fmd_cont <- fmd$controls

# Estimate bandwidths and create density/risk maps
hlam <- LIK.spattemp(fmd_case)
f.breve <- spattemp.density(fmd_case, h=hlam[1], lambda=hlam[2], tlim=c(10, 352))
g.tilde <- bivariate.density(fmd_cont, h0=hlam[1])
rho.breve <- spattemp.risk(f=f.breve, g=g.tilde, tolerate=TRUE)
```

```{r}
# Set up color map
colpal <- spatstat.options("image.colfun")

# Adjust the color map accordingly
fmdcolours <- colourmap(colpal(320), breaks=c(-80, seq(-7,4,length=320)))

# Time range for KDE plot
mytimes <- 10:352

# Save the animation as a GIF
animation::saveGIF(
  for (i in 1:length(mytimes)) {
    # Plot KDE with time selection
    plot(rho.breve, tselect=mytimes[i], type="conditional", col=fmdcolours,
         tol.args=list(levels=c(0.05, 0.0001), lty=2:1, drawlabels=FALSE),
         sleep=0, override.par=FALSE, fix.range=TRUE)
    title(xlab="Easting", ylab="Northing")
  },
  movie.name="fmd_updated.gif", ani.width=640, ani.height=640, interval=1/12
)
```
